---
title: "Project2_Allen_Nicholas"
output: html_document
Subject:    Project 2
Class:      DSCI 512
Section:    1W 24/SP2    
Instructor: Nengbing Tao
date: "2024-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
mtcars <- read_excel("C:/Users/user1/OneDrive/Documents/DSCI 512 Predictive Modelling/Week 2/mtcars.xlsx")
View(mtcars)
```
### 1. Use the lm() function to perform a simple linear regression with the 
### response mpg and the predictor hp.

```{r}
lm.fit = lm(mpg~hp, data=mtcars)
summary(lm.fit)
```

### 2. Is there a relationship between the target mpg and predictor hp?
Since the p-value is far less than 5% (2.723E-09) we can say there is clear 
evidence to reject the NULL hypothesis. This means that there is evidence of 
a relationship between the predictor ``` hp ``` and the target ``` mpg ```.


### 3. How strong is the relationship between the response and predictor?
Accuracy measures can help us determine how strong the relationship is. 
R-squared is just over 60% for us. This indicates only a mild relationship vice
a strong relationship.


### 4. Is the relationship between mpg and hp positive or negative?
The relationship is negative as pointed out by the  ``` hp ``` coefficient -0.06

### 5. What is the predicted mpg associated with a horsepower (hp) of 100? 
### What’s the 95% confidence interval for the predicted mpg?
The predicted mpg for 100 hp should be ``` 29.297928+(-0.064548)*100=22.8431 mpg
```  

The 95% confidence interval is found using 
```{r}
predict(lm.fit, data.frame(hp = 100), interval = "confidence")
```

### 6. Plot the response and the predictor and add the regression line using 
### abline().
```{r}
plot(mtcars$hp, mtcars$mpg)
abline(lm.fit, lwd=3, col="blue")
```

### 7. Perform a multiple linear regression with mpg as the response and the 
### predictors cyl, disp, hp, wt, vs, and gear. Print out the results using
### summary() function.
```{r}
lm.fit_mult = lm(mpg~cyl+disp+hp+wt+vs+gear, data=mtcars)
summary(lm.fit_mult)
```

### 8. Is there a relationship between the predictors and the response?
Since the p-value is far less than 5% (2.568E-12) we can say there is clear 
evidence to reject the NULL hypothesis. This means that there is evidence of 
a relationship between the predictors and the target.

### 9. Which predictors appears to have a statistically significant relationship
### to the response?
To determine which of multiple predictors have a meaningful relationship to the
response, we must look at the individual p values displayed for each predictor.
These are denoted in the ```Pr(>|t|)``` column of the summary output. While 
```hp``` is low (~2%) only ```wt``` is significantly below the 5% threshold
(0.012%) thus we could say ```hp``` has a strong relationship in regressing 
```mpg``` but only ```wt``` is backed up as a statistically significant 
relationship. 

### 10.  Use * symbols to fit linear regression models with interaction effects 
### between hp and wt. Does this interaction appear to be statistically 
### significant?
```{r}
lm.fit_mult2 = lm(mpg~cyl+disp+hp*wt+vs+gear, data=mtcars)
summary(lm.fit_mult2)
```
Now this model seems to prove that ```hp``` does have a statistically 
significant relationship with ```mpg```. This is due to its p value now being 
well below 5% at 0.016% as well as the interaction between ```hp``` and 
```wt``` also having a p value below 5% (0.11%). The p value for ```wt``` 
decreased even further, thus confirming it still has a significant relationship 
with ```mpg```. 
