---
title: "Project7_Allen_Nicholas"
output: html_document
Subject:    Project 7
Class:      DSCI 512
Section:    1W 24/SP2    
Instructor: Nengbing Tao
date: "2024-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 1. Load the dataset CreditCards.csv Download CreditCards.csv into memory.
```{r}
CC <- read.csv("C:/Users/user1/OneDrive/Documents/DSCI 512 Predictive Modelling/Week 7/CreditCards.csv", header=T)
View(CC)
```
### 2. Perform the k-means cluster analysis 
#### * a. Remove the first column: CUST_ID since it doesn’t provide any info for cluster.
```{r}
if (!require("factoextra")) install.packages("factoextra")
if (!require("factoextra")) library(factoextra)
if (!require("cluster")) install.packages("cluster")
if (!require("cluster")) library(cluster)
#seeds allow reproducible results 
set.seed(13)
newCC=CC[-1]
View(newCC)
```


### 2.b Determine the optimal number of clusters. Justify your answer. It may take longer running time since it uses a large dataset.
```{r}
fviz_nbclust(newCC, kmeans, method = "gap_stat")
```


### 2.c Perform k-means clustering using the optimal number of clusters.
```{r}
km.res <- kmeans(newCC, 2, nstart = 25)
```

### 2.d Visualize the clusters in different colors.
```{r}
fviz_cluster(km.res, data = newCC)
```
